\documentclass{article}

\title{Algebraic Topology Homework 5}
\author{Adam Buskirk}

\usepackage{amssymb,amsmath,amsthm}
\usepackage{tikz}
\usepackage[margin=1in]{geometry}

\newtheorem{theorem}[subsection]{Theorem}
\newtheorem{conjecture}[subsection]{Conjecture}
\newtheorem{lemma}[subsection]{Lemma}
\theoremstyle{definition}
\newtheorem{definition}[subsection]{Definition}

\newcommand{\R}{\mathbb{R}}
\newcommand{\N}{\mathbb{N}}
\newcommand{\Q}{\mathbb{Q}}
\newcommand{\Z}{\mathbb{Z}}
\newcommand{\Co}{\mathbb{C}}
\newcommand{\p}[1]{\left(#1\right)}
\newcommand{\sq}[1]{\left[#1\right]}
\newcommand{\set}[1]{\left\{#1\right\}}
\newcommand{\norm}[1]{\left|\left|#1\right|\right|}
% \newcommand{\p}[1]{\left(#1\right)}

\begin{document}
\maketitle

\section{$n$-manifolds}
\begin{theorem}
An $n$-manifold has a countable basis of precompact coordinate balls.
\end{theorem}
\begin{proof}
Suppose $X$ is an $n$-manifold. Since $X$ is a topological manifold, 
$X$ is locally compact (Proposition 4.64). By Proposition 4.63, each
point of $X$ has a precompact neighborhood. Since manifolds are
second-countable, $X$ has a countable basis.
\end{proof}

\section{Compactness implies paracompactness}
\begin{theorem}
Compactness implies paracompactness.
\end{theorem}
\begin{proof}
Suppose a space $X$ is compact. Then for any cover $\mathcal{B}
=\set{B_i, i\in I}$ over $X$,
then by compactness we have a finite subcover 
$\mathcal{B}'=\set{B_i, I\in J} \subseteq \mathcal{B}$. Since each 
$B_i \in \mathcal{B}'$ is an open subset of $B_i \in \mathcal{B}$, 
$\mathcal{B}'$ is an open refinement of $\mathcal{B}$. Furthermore, since
$\mathcal{B}'$ is finite, it must be locally finite. Thus
every open cover of $X$ admits a locally finite open refinement, so $X$
is paracompact.
\end{proof}

\end{document}
